
```{r}
model = "model
{
  # a is in metres
  log_a ~ dnorm(2, 1/2^2)
  a <- exp(log_a)
  
  # b is in metres per kilogram
  log_b ~ dnorm(-2, 1/2^2)
  b <- exp(log_b)
  
  # c is in metres per square kilogram, it's a bit harder
  # to get a handle on so I'm going wider here
  c ~ dt(0, 1, 1)
  # Sigma is in metres
  log_sigma ~ dnorm(2, 1/3^2)
  sigma <- exp(log_sigma)
  
  # Nu (shape parameter of t distribution) is dimensionless
  log_nu ~ dunif(0, 5)
  nu <- exp(log_nu)
  
  for(i in 1:length(y))
  {
    mu[i] <- a + b*(x[i] - mean(x)) + c*(x[i] - mean(x))^2
    y[i] ~ dt(mu[i], 1/sigma^2, nu)
  }
}
"
```

```{r}
data = data.frame(x = c(37.5, 51.5, 61.3, 61.3, 63.6, 66.1,
70, 92.7, 90.5, 90.5, 94.8, 97, 97, 97, 102, 102, 103.6, 100.4,
108.4, 114, 115.3, 114.9, 114.7, 123.6, 125.8, 119.1, 118.9,
141.1), y = c(6.4, 10.2, 12.4, 13, 13.2, 13, 12.7, 13.9, 15.5,
15.8, 15.8, 16.8, 17.1, 17.8, 14.8, 15.5, 16.1, 16.2, 17.9, 15.9,
15.8, 16.7, 17.6, 16.8, 17, 18.2, 19.2, 18.6))

# variables to monitor
variable_names = c("a", "b", "c", "sigma", "nu")

# How many proper steps?
steps = 100000

# Thinning?
thin = 10

# Import the rjags library
library("rjags")

# Create a JAGS model object
jm = jags.model(textConnection(model), data)

# Do some MCMC
results = coda.samples(jm, variable_names, steps, thin=thin)

# Extract chains as data frame
results = as.data.frame(results[[1]])
```

### 1a) Plot the data and get acquainted with it. Read the model and make sure you understand
it.

```{r}
plot(data$x, data$y)
```

Looks mostly linear, slight quadratic trend.

### 1b) Plot many regression curves through the data using parameter values from the MCMC
output, to provide a visualisation of the uncertainty of the fit.

```{r}
{
plot(data$x, data$y)

# plot lines
for (i in 1:1000) {
  lines(x = data$x, y = results$a[i] + results$b[i]*(data$x - mean(data$x)) + results$c[i]*(data$x - mean(data$x))^2, col = rgb(0, 0, 1, alpha = 0.05))
}
}
```
### 1c) 

```{r}
model = "model
{
  # a is in metres
  log_a ~ dnorm(2, 1/2^2)
  a <- exp(log_a)
  
  # b is in metres per kilogram
  log_b ~ dnorm(-2, 1/2^2)
  b <- exp(log_b)
  
  # c is in metres per square kilogram, it's a bit harder
  # to get a handle on so I'm going wider here
  c ~ dt(0, 1, 1)
  # Sigma is in metres
  log_sigma ~ dnorm(2, 1/3^2)
  sigma <- exp(log_sigma)
  
  # Nu (shape parameter of t distribution) is dimensionless
  log_nu ~ dunif(0, 5)
  nu <- exp(log_nu)
  
  for(i in 1:length(y))
  {
    mu[i] <- a + b*(x[i] - mean(x)) + c*(x[i] - mean(x))^2
    y[i] ~ dt(mu[i], 1/sigma^2, nu)
  }
  
  # point prediction nodes
  x_new <- 40
  mu_new <- a + b*(x_new - mean(x)) + c*(x_new - mean(x))^2
  y_new ~ dt(mu_new, 1/sigma^2, nu)
}
"

# variables to monitor
variable_names = c("a", "b", "c", "sigma", "nu", "y_new")

# How many proper steps?
steps = 100000

# Thinning?
thin = 10

# Import the rjags library
library("rjags")

# Create a JAGS model object
jm = jags.model(textConnection(model), data)

# Do some MCMC
results = coda.samples(jm, variable_names, steps, thin=thin)

# Extract chains as data frame
results = as.data.frame(results[[1]])

hist(results$y_new, breaks = 100)
```

```{r}
sorted = sort(results$y_new)
l = length(sorted)
c(sorted[0.025*l], sorted[0.975*l])
```


Given the data, there is a 95% probability that the woman who lifts 40 kg can shotput between
5.5 and 10.9 metres.

### 1d) A slightly harder prediction: Add node(s) to the model to predict the shot-put
distance for a woman whose weightlifting ability we do not know, but to which we assign
a Normal(40, 102)T (0, ) prior2. Compute a 95% credible interval for it. Sanity check: It
should be wider than the interval in (c), for obvious reasons.

```{r}
model = "model
{
  # a is in metres
  log_a ~ dnorm(2, 1/2^2)
  a <- exp(log_a)
  
  # b is in metres per kilogram
  log_b ~ dnorm(-2, 1/2^2)
  b <- exp(log_b)
  
  # c is in metres per square kilogram, it's a bit harder
  # to get a handle on so I'm going wider here
  c ~ dt(0, 1, 1)
  # Sigma is in metres
  log_sigma ~ dnorm(2, 1/3^2)
  sigma <- exp(log_sigma)
  
  # Nu (shape parameter of t distribution) is dimensionless
  log_nu ~ dunif(0, 5)
  nu <- exp(log_nu)
  
  for(i in 1:length(y))
  {
    mu[i] <- a + b*(x[i] - mean(x)) + c*(x[i] - mean(x))^2
    y[i] ~ dt(mu[i], 1/sigma^2, nu)
  }
  
  # point prediction nodes
  sigma_new <- 10
  x_new ~ dnorm(40, 1/sigma_new^2)T(0,)
  mu_new <- a + b*(x_new - mean(x)) + c*(x_new - mean(x))^2
  y_new ~ dt(mu_new, 1/sigma^2, nu)
}
"

# variables to monitor
variable_names = c("a", "b", "c", "sigma", "nu", "y_new")

# How many proper steps?
steps = 100000

# Thinning?
thin = 10

# Import the rjags library
library("rjags")

# Create a JAGS model object
jm = jags.model(textConnection(model), data)

# Do some MCMC
results = coda.samples(jm, variable_names, steps, thin=thin)

# Extract chains as data frame
results = as.data.frame(results[[1]])

hist(results$y_new, breaks = 100)
```

```{r}
sorted = sort(results$y_new)
l = length(sorted)
c(sorted[0.025*l], sorted[0.975*l])
```

Credible interval appears quite a lot wider when there's uncertainty about the weightlifting ability of the woman.


