```{r setup, include=FALSE}
model_1 = "model
{
  # Regression coefficients
  beta0 ~ dnorm(0, 1/1000^2)
  beta1 ~ dnorm(0, 1/1000^2)
  
  # Noise standard deviation
  log_sigma ~ dnorm(0, 1/10^2)
  sigma <- exp(log_sigma)
  
  # Data
  for(i in 1:length(y))
  {
    mu[i] <- beta0 + beta1*x[i]
    y[i] ~ dnorm(mu[i], 1/sigma^2)
  }
}
"

# A dataset
data = read.csv("jumping.csv")

# Variables to monitor
variable_names = c("beta0", "beta1")
```

### 1a) Plot the data so you know what you’re looking at. Also, read the model closely and make
sure you understand its assumptions.

```{r}
plot(data$x, data$y, xlab = "x", ylab = "y")
```

Appear roughly linear, slightly skewed data.

### 1b) Run the regression model and plot the joint posterior of the two coefficients β0 and β1.
You should see a fairly strong dependence. Can you explain why this occurs? Hint: It’s
something to do with the x-values, and it might be clearer after you’ve done (c).

```{r}
# How many proper steps?
steps = 100000

# Thinning?
thin = 10

# Import the rjags library
library("rjags")

# Create a JAGS model object
jm = jags.model(textConnection(model_1), data)

# Do some MCMC
results = coda.samples(jm, variable_names, steps, thin=thin)

# Extract chains as data frame
results = as.data.frame(results[[1]])

pairs(results)
```
Strong dependance between beta0 and beta1

### 1c) Strong dependence can cause trouble for JAGS (though it wasn’t too bad here). The
strong dependence can be removed by re-defining the covariate(s) as differences from their
arithmetic mean _____. In this case, the change is:

mu[i] <- beta0 + beta1*(x[i] - mean(x))

Make this change to the model and run it again. Plot the joint posterior for β0 and β1
after the change.

```{r}
model_2 = "model
{
  # Regression coefficients
  beta0 ~ dnorm(0, 1/1000^2)
  beta1 ~ dnorm(0, 1/1000^2)
  
  # Noise standard deviation
  log_sigma ~ dnorm(0, 1/10^2)
  sigma <- exp(log_sigma)
  
  # Data
  for(i in 1:length(y))
  {
    mu[i] <- beta0 + beta1*(x[i] - mean(x))
    y[i] ~ dnorm(mu[i], 1/sigma^2)
  }
}
"

# Create a JAGS model object
jm = jags.model(textConnection(model_2), data)

# Do some MCMC
results = coda.samples(jm, variable_names, steps, thin=thin)

# Extract chains as data frame
results = as.data.frame(results[[1]])

pairs(results)
```

Beta0 and beta1 appear appropriately normally distributed

### 1d) 

Pen and paper, need to equate coefficients to find relationship between beta0 and beta1.

### 1e)

Okay do this later haha.
