### Class Example

#### Grid method for Uniform Prior

```{r}
theta = seq(0.001, 0.999, by=0.001)
lik = dbinom(13, size=16, theta)

plot(theta, lik, type="l")
```

Uniform Prior Bayes Factor

```{r}
prior_uniform = rep(1, length(theta))
plot(theta, prior_uniform, type="l")
```

```{r}
Z0 = dbinom(13, size=16, prob=0.5)
Z1 = 0.001*sum(prior_uniform * lik)

bayes_factor_uniform = Z1/Z0
bayes_factor_uniform
```

6.9x more likely under the alternative hypothesis than the null hypothesis that theta = 0.5.

Jeffrey's Prior Bayes Factor

```{r}
prior_jeff = dbeta(theta, 0.5, 0.5)

plot(theta, prior_jeff, type="l")
```

```{r}
Z0 = dbinom(13, size=16, prob=0.5)
Z2 = 0.001*sum(prior_jeff * lik)

bayes_factor_jeff = Z2/Z0
bayes_factor_jeff
```

Similar Prior Bayes Factor

```{r}
prior_similar = dbeta(theta, 5, 4)

plot(theta, prior_similar, type="l")
```

```{r}
Z0 = dbinom(13, size=16, prob=0.5)
Z3 = 0.001*sum(prior_similar * lik)

bayes_factor_similar = Z3/Z0
bayes_factor_similar
```



### Tutorial

```{r}
data = list(x=c(1.5, 2.5, 1.7), N=3)
```


Consider two hypotheses about the value of μ. Let H0 be the hypothesis that μ = 0, and let H1
be the hypothesis that μ ̸ = 0 but that it has a uniform prior between -10 and 10.

1a) Calculate the likelihood for H0.

```{r}
# priors
mu = seq(-10, 10, 0.01)
prior = dunif(mu, -10, 10)
prior = prior / sum(prior)

plot(mu, prior)

# compute likelihood for H0
lik_h0 = dnorm(data$x[1], 0, 1) * dnorm(data$x[2], 0, 1) * dnorm(data$x[3], 0, 1) 
marg_lik_h0 = lik_h0
```

1b) Calculate the likelihood for H1 as a function of μ and hence find the marginal likelihood for H1.

```{r}
# compute likelihood for H1 as function of mu
lik_h1 = dnorm(data$x[1], mu, 1) * dnorm(data$x[2], mu, 1) * dnorm(data$x[3], mu, 1)

plot(mu, lik_h1)

# compute marginal likelihood for H1 as function of mu
marg_lik_h1 = sum(prior * lik_h1)

marg_lik_h1
```

1c) Compute (i) the Bayes Factor; (ii) the posterior odds ratio assuming a prior odds ratio of 1; and (iii) the posterior probabilities of the two hypotheses.

```{r}
bayes_factor = marg_lik_h0 / marg_lik_h1

prior_odds = 1

posterior_odds = prior_odds * bayes_factor

posterior_probs_h0 = prior * lik_h0 / (marg_lik_h0 + marg_lik_h1)
plot(mu , posterior_probs_h0)
```
```{r}
posterior_probs_h1 = prior * lik_h1 / (marg_lik_h0 + marg_lik_h1)
plot(mu , posterior_probs_h1)
```

1d) What happens to the marginal likelihood as you widen the prior for μ under H1?

It will get smaller.

