---
title: "Untitled"
output: html_document
date: "2023-10-17"
---

The USA Junior Chess Olympics Research: Developing Memory and Verbal Reasoning” (New
Horizons for Learning, April 2001; (available at www.newhorizons.org) describes a study in
which sixth-grade students who had not previously played chess participated in a program in
which they took chess lessons and played chess daily for 9 months. Each student took a memory
test (the Test of Cognitive Skills) before starting the chess program and again at the end of
the 9-month period. 

The data below has the differences in scores after the chess program and
before, for 12 students.

```{r}
data = list(x=c(340, 180, 210, 100, 100, 225, 90,
225, 240, 55, -35, 5), N=12)
```

2a) Do a classical t-test using R (this should be very easy). Write down the p-value, and write
down the notation for what this probability is (i.e., it’s the probability of what conditional
on what)

```{r}
t.test(data$x)
```

Alternatively

```{r}
t_test = mean(data$x) / (sd(data$x) / sqrt(data$N))
t_test
```


The p-value is the probability of a more extreme value of the test statistic if H0 is true.

t = x_mean / s / sqrt(N)

where s is the sample standard deviation (the version with N − 1 in the denominator).

The observed value of t was 4.564. The p-value is the probability of a more extreme value than this, under H0:
  P (t ≥ 4.564 or t ≤ −4.564 | H0)

2b) Do Bayesian model comparison of H0 vs. H1. You can compute marginal likelihoods using
either Nested Sampling or another method if you prefer. For H1, use a Uniform(-1000000,
1000000) prior for μ. Calculate the posterior odds ratio assuming a prior odds ratio of 1,
and write down the probability notation for what this quantity is.

H0 : mu = 0
H1 : mu /= 0

Todo: setup the file such that we run all the files from here

```{r}
source("ns_tutorial/nested-sampling-tutorial.R")

results = read.csv("ns-output.csv")
```

